How to select predictive models for decision-making or causal inference?

### 2024-09-30 response_to_reviewer.md

Let me begin with positive comments.  I think the paper is very important because it describes a sophisticated methodology for experimental examination of causal models.
Despite some attempts this is still lacking in literature.  I especially like the simulation scheme which uses different features for data generation and learning models.  Moreover thorough experimental evaluations of causal models are not very frequent making the paper important.  Code is available on Github which is a big plus.

I am however not sure about the overall conclusions, which are that R-risk is always the best.  Performance of risk functions may depend on several factors and the overall best performance might be a result of bias in available data.  The only factor taken into account by the authors is overlap.  However, only one type of lack of overlap is tested in experiments, based on shifted Gaussians.  This will likely lead to violation of the positivity assumption.  How about different scenarios, e.g.: perfect overlap but different treatment/control distributions?  Confounding need not necessarily lead to lack of overlap.  Another possible factor is the ratio of causal effect to background responses.  It is known that it may affect relative performance of HTE estimators.

Overall: I am no fully convinced that R-risk is always the best.  The answer could be dependent on other factors and I hope the authors would explore more scenarios.

> We thank the reviewer for this remark and the suggestion to explore more diverse simulation scenarii. Exploring the sensitivity of our result on the R-risk appears to us particularly necessary. Thus, we ran a dedicated experiment focusing on the ratio of causal effect to background response.

> We did not explore more sophisticated treatment/control distributions for our simulation. Building these new data generation processes would imply a big modification of our simulation. We rely on the easy inversion of the distribution functions of the gaussians (p(X|A=a)) to analytically compute the true propensity score. In addition, we think that the semi-simulated datasets already provide some interesting argument in favor of r-risk with more complex treatment/control covariate distributions. However, we mentionned in the discussion the need to explore more diverse data generation processes for simulations.

### 2024-09-30 how_to_select_latex.tex
reviewer response : wip on the last point

\begin{frontmatter}
    \maketitle
    \begin{abstract}
        \textbf{Background:}
        We investigate which procedure selects the predictive model most
        trustworthy to reason on the effect of an intervention and support
        decision making.\\
        %As predictive models --\eg from machine learning-- give likely
        %outcomes, they may be used to, a
        %causal-inference task. The increasing complexity of health data has opened
        %the door to a plethora of models, but also the Pandora box of model
        %selection: which of these models yield the most valid causal estimates? 
        \textbf{Methods:}
        We study a large variety of model selection procedures in practical settings:
        finite samples settings and without theoretical assumption of
        \emph{well-specified} models. Beyond standard cross-validation or
        internal validation procedures, we also study elaborate causal risks.
        These build proxies of the causal error using ``nuisance''
        re-weighting to compute it on the observed data. We evaluate whether
        empirically estimated nuisances, which are necessarily noisy,
        add noise to model selection. We compare different metrics for
        causal model selection in an extensive empirical
        study based on a simulation and three healthcare datasets
        based on real covariates. \\
        \textbf{Results:} Among all metrics, the mean squared error,
        classically used to evaluate predictive modes,
        is worse. Re-weighting it with propensity score
        does not bring much improvements. The
        $R\text{-risk}$, which uses as nuisances a model of
        mean outcome and propensity scores, leads to the best performances.
        Nuisance corrections are best estimated with flexible estimators such
        as a super learner.
        \\
        \textbf{Conclusions:} When predictive models are used to
        reason on the effect of an intervention, they must be evaluated
        with different procedures than standard predictive
        settings; using the $R\text{-risk}$ from causal inference.
    \end{abstract}

@@

@@

>>>>

\begin{frontmatter}
    \maketitle
    \begin{abstract}
        \textbf{Background:}
        We investigate which procedure selects the predictive model most
        trustworthy to reason on the effect of an intervention and support
        decision making.\\
        %As predictive models --\eg from machine learning-- give likely
        %outcomes, they may be used to, a
        %causal-inference task. The increasing complexity of health data has opened
        %the door to a plethora of models, but also the Pandora box of model
        %selection: which of these models yield the most valid causal estimates? 
        \textbf{Methods:}
        We study a large variety of model selection procedures in practical settings:
        finite samples settings and without theoretical assumption of
        \emph{well-specified} models. Beyond standard cross-validation or
        internal validation procedures, we also study elaborate causal risks.
        These build proxies of the causal error using ``nuisance''
        re-weighting to compute it on the observed data. We evaluate whether
        empirically estimated nuisances, which are necessarily noisy,
        add noise to model selection. We compare different metrics for
        causal model selection in an extensive empirical
        study based on a simulation and three healthcare datasets
        based on real covariates. \\
        \textbf{Results:} Among all metrics, the mean squared error,
        classically used to evaluate predictive modes,
        is worse. Re-weighting it with propensity score
        does not bring much improvements in most cases. On average, the
        $R\text{-risk}$, which uses as nuisances a model of
        mean outcome and propensity scores, leads to the best performances.
        Nuisance corrections are best estimated with flexible estimators such
        as a super learner.
        \\
        \textbf{Conclusions:} When predictive models are used to
        reason on the effect of an intervention, they must be evaluated
        with different procedures than standard predictive
        settings; using the $R\text{-risk}$ from causal inference.
    \end{abstract}

@@

\paragraph{R-risk is robust to a wide range of effect ratio}

We focused on the

\begin{figure}[!tb]
    % XXX: for submission
    \centering\begin{minipage}{\linewidth}
        \includegraphics[width=\linewidth]{_6_effect_ratio_by_bin_comparaison_kendall_by_Dataset.pdf}
    \end{minipage}

    \caption{\textbf{R-risk is robust to a wide range of effect ratio}: Kendall’s $\tau$ agreement with $\tau$-risk. Strong, medium and Weak Causal effect ratio are the tertiles of the absolute ratio causal effect to baseline response, ranging from 1 to 300. Appendix \ref{apd:effect_ratio} details this simulation.}\label{fig:effect_ratio_influence}
\end{figure}

@@

\subsection{Sensitivity of the results to different ratio of causal effect to baseline response}\label{apd:effect_ratio}

TODO
